---
title: "Experiments with Forest Products Trade Flow Data"
output:
  html_document:
    toc: true
---

```{r, echo=FALSE}
# Set same path for knitr evaluation as for interactive use
library(knitr)
opts_knit$set(root.dir = '../..')
opts_chunk$set(fig.width=10)
```

Load functions and scenario data
```{r warning=FALSE}
library(dplyr)
source("code/func.R")
load("enddata/GFPM_training_scenarios.RDATA")
load("enddata/GFPM_training_scenarios_with_historical.RDATA")
```
Base scenario + other 2 training scenarios, calculated by changing the demand
elasticities by plus or minus 1 standard error, corresponding to a 
confidence intereval of 70%.

Introduction
============
Explore country specific data.
Unfortunately GFPM doesn't simulate bilateral trade.
### Todo
add all elements on one graph grid to after individual plots to help visualiwe differences in scale.

```{r}
country = "United Kingdom"
dtf = subset(trainingScenarios$entity, Country == country & 
                 Scenario=="Base" )
```

### Issues by product

Industrial roundwood decreasing and increasing in the base2011 scenario.
```{r}
ggplot(data=filter(gfpm, Country==country & 
                       Product %in% c("IndRound")& Scenario == "Base2011")) +
    aes(x=Year, y=Volume, color=Element, linetype=Scenario) +
    geom_line() + ylim(0,NA) + 
    ylab(expression(paste("Volume ",m^3))) +
    facet_grid(Element ~ Product, scales="free_y") 
```



## Plot by element

```{r by_element, results='markup'}
# Import
for (element in unique(dtf$Element)){
#     print(head(subset(dtf, Element == element)))
    print(paste("###", element))
    p = ggplot(data=subset(dtf, Element == element)) +
        aes(x=Period, y=Volume, colour=Product, label = Product) +
        ggtitle(paste(country,"-",element)) +
        geom_line() + 
        # Subset last period to print country labels on the curves
        geom_text(data=subset(dtf, Element == element & Period==max(dtf$Period))) +
        xlim(1,5.5)
    print(p)
    #            theme(legend.position = "none") 
}
```


### All on one graph
```{r}
ggplot(data=subset(dtf, Element!="DPrice")) +
    aes(x=Period, y=Volume, colour=Product, label = Product) +
    ggtitle(paste(country)) +
    geom_line() + 
    # Subset last period to print country labels on the curves
    geom_text(data=subset(dtf, Element!="DPrice" & Period==max(dtf$Period))) +
    facet_wrap(~Element) 
```


## Plot by product
Except for Prices
```{r by_product, results='markup'}
# Import
for (product in unique(dtf$Product)){
#     print(head(subset(dtf, Element == element)))
    cat(paste("###", product))
    dtf_product <- subset(dtf, Product == product & Element!="DPrice")
    p = ggplot(data=dtf_product) +
        aes(x=Period, y=Volume, colour=Element, label = Element) +
        ggtitle(paste(country,"-",product)) +
        geom_line() + 
        # Subset last period to print country labels on the curves
        geom_text(data=subset(dtf_product, Period==max(dtf_product$Period))) +
        xlim(1,5.5)
    print(p)
    #            theme(legend.position = "none") 
}
```

### All on one graph
```{r}
ggplot(data=subset(dtf, Element!="DPrice")) +
    aes(x=Period, y=Volume, colour=Element, label = Element) +
    ggtitle(paste(country)) +
    geom_line() + 
    # Subset last period to print country labels on the curves
    geom_text(data=subset(dtf, Element!="DPrice" & Period==max(dtf$Period))) +
    facet_wrap(~Product) +
    xlim(1,6)
```


### Plot by product type
Primary secondary product.


Data summaries
--------------
### What products are demanded, supplied, produced or traded in the country?
```{r eval=FALSE}
whatproducts = function(dtf){unique(dtf$Product)}
a <- group_by(dtf, Element)
per_element <- summarize(a, p = unique(a$Product))
# dlply(dtf, .(Element), whatproducts)[-2]
do(dtf, prod = 1)
# unique(trainingScenarios$entity$Country)
# unique(trainingScenarios$entity$Element)
```

Explore GFPM scenarios by continents
================================
Plots
-----
### Base scenario
```{r sawnwood_base}
plotProdByReg(trainingScenarios, "Sawnwood", "Base")
```

```{r roundwood_base}
plotProdByReg(trainingScenarios, "IndRound", "Base")
```
