Explore TTIP dataset
=====================
```{r, echo=FALSE, warning=FALSE}
# Set same path for knitr evaluation as for interactive use
opts_knit$set(root.dir = '../..')
```

Load a dataset containing GFPM scenarios
```{r}
library(plyr)
library(xtable)
library(ggplot2)
load("enddata/GFPM_Output_TTIP_with_sensitivity.RDATA")
```


Sensitivity analysis 
---------------------
Plot scenarios changes

```{r}
product = "Sawnwood"
element = "Demand"
aggregates=subset(allScenarios$aggregates, Scenario !="LowTTIP")
dtf = subset(aggregates, Product==product)


p = ggplot(data = subset(dtf,Element == element&Scenario!="HighTTIP")) +
    aes(x=Period, y=Volume, colour=GFPM_REG) +
    geom_line(aes(linetype = Scenario)) + 
#     geom_point(aes(shape=Scenario)) +
#     geom_point(data = subset(dtf,Element == element&Scenario=="HighTTIP"),
#                shape=15, size=4) +
    geom_point(data = subset(dtf,Element == element&Scenario=="HighTTIP"),
               aes(shape=Scenario), size=4) +
    ggtitle(paste(product, element)) +
    theme(legend.position = "bottom") 
print(p)
```

### Make a similar  graph for US - EU28 - Rest of the world


Comparison with Felbermayr 2013
-------------------------------
Felbermayr 2013 reports a simulation of US and Germany's export values by 2025 for 3 forest related sectors: "forestry", "wood products" and "paper". 

What are the export products?
```{r}
unique(allScenarios$entity$Product[allScenarios$entity$Element=="Export"]) 
``` 

We will aggregate the products as such:
* __Forestry__ :  IndRound, 
* __Wood Products__ :  "Fuelwood" "Sawnwood"   "Plywood"    "ParticleB"  "FiberB"  
* __Paper Products__ : "MechPlp"  "ChemPlp"    "OthFbrPlp"  "WastePaper" "Newsprint"  "PWPaper"    "OthPaper"  
    
US and Germany's exports in forest products in 2025 (period 4 in our simulation). We are looking at percentage changes in volume between the base and the HighTTIP scenarios.
```{r results='asis'}
USDEexp = subset(allScenarios$entity,
                 Country%in% c("United States of America","Germany") &
                     Element=="Export" & 
                     Period == 4 &
                     Scenario %in% c("Base", "HighTTIP"),
                 select=c(Scenario, Period, Product, Country, Volume))
USDEexp = reshape(USDEexp,
                  idvar = c("Period", "Product", "Country"),
                  timevar = "Scenario", times=c("Base", "HighTTIP"),
                  direction="wide")
USDEexp = transform(USDEexp, expchange = round((Volume.HighTTIP - Volume.Base)/ Volume.Base *100,2))

USDEexp_for_JB = reshape(subset(USDEexp), 
                  idvar = c("Period", "Product"),
                  timevar = "Country",
                  direction="wide")

USDEexp = reshape(subset(USDEexp,select=-c(Volume.Base,Volume.HighTTIP)), 
                  idvar = c("Period", "Product"),
                  timevar = "Country",
                  direction="wide")
names(USDEexp) = c("Period", "Product", "US exports %", "Germany exports %")


print(xtable(USDEexp, caption = "Percentage change in exports between the 
             base and HighTTIP scenarios on period 4 (2015)"), 
      type = "html", include.rownames = FALSE) 
USDEexp$ProductAgg[USDEexp$Product %in%
                       c("Fuelwood", "Sawnwood", "Plywood", "ParticleB", "FiberB")] = "Wood Products"
USDEexp$ProductAgg[USDEexp$Product == "IndRound"] = "Forestry"
# USDEexp$ProductAgg[USDEexp$Product %in% c("MechPlp", "ChemPlp", "OthFbrPlp", "WastePaper",
#                                           "Newsprint", "PWPaper", "OthPaper") = "Paper Products"
```
