---
title: "Explore GFPM data"
author: "Paul Rougieux"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---


```{r warning=FALSE, message=FALSE}
library(GFPMoutput)
library(knitr)
library(dplyr)
library(ggplot2)
library(scales)
```

```{r echo=FALSE}
# Github flavoured markdown for tables
# http://stackoverflow.com/questions/28687826/how-to-generate-a-github-flavoured-markdown-file-using-knitr
options(knitr.table.format = 'markdown')
```


```{r, echo=FALSE}
# Set same path for knitr evaluation as for interactive use
opts_knit$set(root.dir = '../..')
opts_chunk$set(fig.width=10)
```

# Load scenario data
If you created a scenario with the `load_and_clean_gfpm_data()` function,  your scenario will be stored in the `enddata` folder.
The data object used for demonstration purposes is called `trainingscenarios`.
It is a list of data frames. 
```{r}
basescenario <- readRDS("enddata/basescenario.rds")
trainingscenarios <- readRDS("enddata/GFPM_training_scenarios.rds")
```
To bind several scenarios together in one list,
use the `bindscenarios()` function.


## What is the structure of "trainingscenarios"
`trainingscenarios` is a list of data frames:

 * scenario data frame contains scenario names and file names
 * entity data frame contains the Supply, DPrice, Import, Demand, Export and Production data at the country level
 * aggregates contains the Supply, DPrice, Import, Demand, Export and Production at the regional level
 * worldPrices contains the world prices
 
```{r}
str(trainingscenarios)
```

You can extract individual data frames with the command:
```{r}
countrydata <- trainingscenarios$entity 
regiondata <- trainingscenarios$aggregates 
```


## How many periods are there per scenario?
```{r}
trainingscenarios$aggregates %>% group_by(Scenario) %>% 
    summarise(number_of_periods = max(Period)) %>% kable()
# Remove periods for base2011 so that all scenarios have the same periods for plotting
trainingscenarios$aggregates <- trainingscenarios$aggregates %>%
    filter(Period <= 5)
trainingscenarios$entity <- trainingscenarios$entity %>%
    filter(Period <= 5)
```


## What products are demanded, supplied, produced or traded ?

* IndRound is supplied and traded but not demanded, it's a primary product.
* OthFbrPlp and WastePaper are supplied and traded but not demanded, they are primary products.
* Fuelwood is both supplied and demanded it's both a primary and a final product.
* Fuelwood is not the outcome of a production process.

```{r}
trainingscenarios$entity %>% group_by(Element) %>%
    summarise(Products = paste(unique(Product),collapse=", ")) %>% kable()
```

## What countries are in the simulation?
```{r countriesingfpm, results='asis'}
countries <- sort(unique(trainingscenarios$entity$Country))
# Convert to utf8 because of character encoding issue
cat(iconv(countries,"latin1","UTF-8"),sep= ", ")
```



# Plots
## Plot by region
```{r sawnwood_base}
plotprodbyreg(trainingscenarios, "Sawnwood", "Base")
```

```{r roundwood_base}
plotprodbyreg(trainingscenarios, "IndRound", "Base")
```


## Plot by country 
Sample plot for Sanwood demand in the base scenario, in France and Germany.
```{r countryplotbase}
plotprodbycounty(trainingscenarios, "Sawnwood", "Base",
                 c("France", "Germany"))
```

The above function generates a plot based on the given arguments.
For greater flexibility you are encouraged to read about 
[ggplot2 package](ggplot2.org). 
Here is an example plot showing 
demand, export, import and production of selected panel products 
in the United States in the base scenario.

```{r countryplotbaseallproducts, eval=TRUE}
usademand <- trainingscenarios$entity %>% 
    filter(Country == "United States of America" & 
               Element %in% c("Demand", "Production",
                              "Import", "Export") &
               Product %in% c("ParticleB", "FiberB","Plywood") &
               Scenario == "Base")

ggplot(data=usademand) +
    aes(x=Period, y=Volume, colour=Product) +
    geom_line() + 
    theme(legend.position = "bottom") +
    facet_grid(~ Element) 
```
Example of a description for this graph. 
The base scenario simulated 
a decrease in particle board production 
and an increase in demand over the 5 periods. 
The increased particle board demand was compensated by increases in imports. On the other hand fibre board and plywood import decreased over the period.

## Compare scenarios
Two training scenarios where calculated by changing the demand
elasticities by plus or minus 1 standard error, corresponding to a 
confidence intereval of 70%.
```{r sawnwood_demand}
sawnwood <- subset(trainingscenarios$aggregates, Product=="Sawnwood"& Element=="Demand")
ggplot(data = sawnwood) +
    aes(x = Period, y = Volume, colour = GFPM_REG, linetype = Scenario) +
    geom_line() + ggtitle("Sawnwood Demand") +
    theme(legend.position = "bottom")
```

Compare for all products
```{r compare_demand_scenarios}
paperproducts <- subset(trainingscenarios$aggregates, Element=="Demand" & 
                  ! Product %in% c("MechPlp", "ChemPlp", "WastePaper"))
ggplot(data = paperproducts) +
    aes(x = Period, y = Volume, colour = GFPM_REG, linetype = Scenario) +
    geom_line() + ggtitle("Demand") +
    theme(legend.position = "bottom") + facet_wrap(~Product, scales="free_y")
```

