---
output: 
  html_document: 
    keep_md: yes
---
Tutorial explaining how to load and clean GFPM data
===================================
```{r, echo=FALSE, warning=FALSE}
library(plyr)
library(knitr)
# Set same path for knitr evaluation as for interactive use
opts_knit$set(root.dir = '../..')
```

Load the package
```{r}
library(GFPMoutput)
```

### Import GFPM data directly after a simulation
This command loads GFPM data in the form of a list of
data frames. It saves the list in R data format
under the folder  ./enddata.
Give a scenario name of your choice.
```{r load_and_clean, eval=FALSE}
load_and_clean_gfpm_data(scenario_name = "your_scenario_name", 
                         pelps_folder = "C:/PELPS/pelps/") 
load_and_clean_gfpm_data(scenario_name = "your_scenario_name",
                         compression="bzip2")
# Test on development system
load_and_clean_gfpm_data(scenario_name = "basescenario3", 
                         pelps_folder = "rawdata/PELPS 105Base") 

# The cleaned data can be loaded with 
load("enddata/your_scenario_name.RDATA")
# It has the form of a large list containing several data frames
# You can see the begining of each data frame with the command
lapply(scenario,head)
```
If you are not interested in the internal workings of load.R and clean.R, 
and if you only want to analyse results from one scenario,
you can skip the rest of this document and jump to tutorial/explore.

However if you want to analyse several scenarios combined, keep on reading.
Below we demonstrate the use of the functions `savePELPSToRdata()` and `clean()`
to read PELPS data tables.


### Copy GFPM data in c:\PELPS\pelps with a function
If the GFPM simulation and R analysis run on two different computers,
it can be convenient to load the data from a named folder 
or archive. To this end, on the GFPM machine, 
GFPM simulation results are stored in plain text
format in ".DAT" files under the `C:\PELPS\pelps` folder. 
After each scenario run, copy this folder into another folder or  archive. Copying the PELPS folder can also be done by hand if R is not installed on that machine. See section below. 

In this example, we give the name "dummy" to the scenario
```{r eval=FALSE}
copy_pelps_folder(scenario_name = "dummy")
```

To save space, data can be compressed
```{r eval=FALSE}
copy_pelps_folder(scenario_name = "dummy", compression="bzip2")
```

Unfortunately zip compression is not available in writting, it can only be read by R.
To use zip compression, follow the method "by hand"" below.

### Copy GFPM data in c:\PELPS\pelps by hand
 1. Rename the folder "C:\PELPS\PELPS" to a name of your choice.
 For example "my_scenario_name". 
 2. Copy the renamed PELPS folder to the \rawdata folder.
   Optionally, you can compress this PELPS folder as a zip file: "my_scenario_name.zip"
   and copy it to the rawdata folder.


Load a scenario in R and save it to .RDATA
-----------------------------------------
Beware that for compressed files, the internal folder in the 
zip archive should have the same name as the zip archive itself.

* Input: Raw PELPS text files (.DAT) stored in a folder or .zip archive
* Output: Raw data.frames stored in a .RDATA file 

Save the base scenario without compression to a RDATA file
```{r eval=FALSE}
savePELPSToRdata("base")
```

Save the base scenario from a bzip2 archive to a RDATA file
```{r eval=FALSE}
savePELPSToRdata("base", compressed="bzip2")
```


Save the base scenario from a zip archive to a RDATA file
```{r eval=FALSE}
savePELPSToRdata("PELPS 105Base", "zip")
```

List Zip archives available
```{r}
list.files("rawdata", ".zip", full.names = TRUE) 
```


List Raw RDATA files available
```{r}
list.files("rawdata", ".RDATA", full.names = TRUE) 
```

Clean a scenario
----------------
The archive is transformed by adding column titles, 
and translating product and country codes into product and country names. 
available for analysis with R.
* Input: Raw data.frames stored in a .RDATA file
* Output: cleaned data.frames stored in a .RDATA file

```{r eval=FALSE}
baseScenario = clean("PELPS 105Base.RDATA", "Base")
```
The `baseScenario` object is a list of dataframe. 
The command `str(baseScenario)` gives details about the structure and content of this data object. 
Now you can explore the dataset in various ways. See the ./docs folder.


Working examples of clean functions
-----------------------------------
### Load raw PELPS tables from a .RDATA file
```{r}
print(load("rawdata/PELPS 105Base.RDATA"))
llply(PELPS,head)
```

### Example using the function splittrade 
```{r}
PELPS2 = splittrade(PELPS)
lapply(PELPS[c("trade")],nrow) #number of rows in PELPS
lapply(PELPS2[c("trade", "import", "export")],nrow) #number of rows in PELPS2
```

### Example using the function reshapeLong
```{r}
head(PELPS$demand)
demand = reshapeLong(PELPS$demand, "Demand")
head(demand)
```

### Example using the function add product and country
```{r}
demand = addProductAndCountry(demand)
head(demand)
```

### Example using reshapeLong and addProductAndCountry on World price
```{r}
wp = reshapeLong(PELPS$worldPrice, "World_Price")
head(wp)
wp = addProductAndCountry(wp)
wp$World_Price = wp$Volume
wp = wp[,c("Product_Code", "Product", "Period", "World_Price")]
head(wp)
```
